---
title: "Final Assignment"
author: "Viva/Yaohan"
date: "`r Sys.Date()`"
output:
  html_document:
    keep_md: yes
    toc: yes
    theme: flatly
    toc_float: yes
    code_folding: hide
    number_sections: no
  pdf_document:
    toc: yes
---
<style>
.kable thead tr th, .table thead tr th {
  text-align: left !important;}
table.kable, table.table {
  width: 100% !important;}
  body {
  line-height: 1.6;
  font-size: 16px
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  dpi = 150)
options(scipen = 999)

if(!require(pacman)){install.packages("pacman"); library(pacman)}
p_load(tidyverse, sf, units, nngeo, tmap, tinytex, kableExtra, janitor, classInt,
       patchwork, here, tidycensus, lwgeom, lsr, haven, descr, RColorBrewer, stargazer,
       car, FNN, GGally, MASS, ISLR2, spdep, caret, ckanr, grid, gridExtra, ggcorrplot,
       jtools, broom, tufte, rmarkdown, corrr, RSocrata, viridis, spatstat, raster, knitr,
       arm, mlogit, readxl, lubridate)

library(conflicted)
conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::recode)
```

# Data Wrangling

1. Transit Data
- Transit Ridership
MBTA Rail Ridership by Time Period, Season, Route/Line, and Stop: https://mbta-massdot.opendata.arcgis.com/datasets/9fea7862501f4c59ab2c001ad6ae19b0_0/explore
The data is for Fall2019, Fall2018, and Fall2017. Since Fall2018 only has data for September, so we only use ridership data for Fall2019 and Fall2017.

- Transit Station
MBTA Rapid Transit: https://www.mass.gov/info-details/massgis-data-mbta-rapid-transit#downloads-
Name and location for each transit stop. Combined with ridership data.

- Definition for "Fall" in Boston
Important Work Restrictions for massDOT: https://www.mass.gov/doc/controls-guidelines-attachment-a-important-work-restrictions/download
The Fall season in Boston is from August.15 to November.1, so we use bike data for August, September and October.

2. Bike Data
- Bike ridership
Blue Bikes Comprehensive Trip Histories: https://s3.amazonaws.com/hubway-data/index.html

- Bike Station
Bluebikes Stations: https://bluebikes.com/system-data

3. Demographic Data

4. Other Data
```{r load_data, message = FALSE, warning = FALSE}
## transit station data

```


## R Markdown


