---
title: "Final Assignment"
author: "Viva/Yaohan"
date: "`r Sys.Date()`"
output:
  html_document:
    keep_md: yes
    toc: yes
    theme: flatly
    toc_float: yes
    code_folding: hide
    number_sections: no
  pdf_document:
    toc: yes
---

<style>
.kable thead tr th, .table thead tr th {
  text-align: left !important;}
table.kable, table.table {
  width: 100% !important;}
  body {
  line-height: 1.6;
  font-size: 16px
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  dpi = 150)
options(scipen = 999)

if(!require(pacman)){install.packages("pacman"); library(pacman)}
p_load(tidyverse, sf, units, nngeo, tmap, tinytex, kableExtra, janitor, classInt,
       patchwork, here, tidycensus, lwgeom, lsr, haven, descr, RColorBrewer, stargazer,
       car, FNN, GGally, MASS, ISLR2, spdep, caret, ckanr, grid, gridExtra, ggcorrplot,
       jtools, broom, tufte, rmarkdown, corrr, RSocrata, viridis, spatstat, raster, knitr,
       arm, mlogit, readxl, lubridate)

library(conflicted)
conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::recode)
```

# Data Preparation

1. Transit Data
- Transit Ridership
MBTA Rail Ridership by Time Period, Season, Route/Line, and Stop: https://mbta-massdot.opendata.arcgis.com/search?tags=ridership
The data is for Fall2019, Fall2018, and Fall2017. Since Fall2018 only has data for September, so we only use ridership data for Fall2019 and Fall2017.

- Transit Station
MBTA Rapid Transit: https://www.mass.gov/info-details/massgis-data-mbta-rapid-transit#downloads-
Name and location for each transit stop； Combined with ridership data.

- Definition for "Fall" in MassDOT
Important Work Restrictions for massDOT: https://www.mass.gov/doc/controls-guidelines-attachment-a-important-work-restrictions/download
The Fall season defined by MassDOT is from August.15 to November.1, so we use bike data for August, September and October.

2. Bike Data
- Bike ridership
Blue Bikes Comprehensive Trip Histories: https://s3.amazonaws.com/hubway-data/index.html
2017: August, September, October
2019: August, September, October

- Bike Station
Bluebikes Stations: https://bluebikes.com/system-data

3. Demographic Data
https://www.bostonplans.org/3d-data-maps/resources
https://data.census.gov/

4. Other Data
- MBTA Boundary
MBTA Core Service Area: https://mbta-massdot.opendata.arcgis.com/search?q=boundary

## Import Datasets

```{r load_data, message = FALSE, warning = FALSE}
# Transit Data
## Transit Ridership
tran_rider <- read_csv(here::here("data/final/Transit_Ridership.csv"))
## Transit Station
tran_line <- st_read(here::here("data/final/mbta_rapid_transit/MBTA_ARC.shp"))
tran_stop <- st_read(here::here("data/final/mbta_rapid_transit/MBTA_NODE.shp"))

# Bike Data
## Bike Ridership
bike_rider_17_aug <- read_csv(here::here("data/final/201708-hubway-tripdata.csv"))
bike_rider_17_sep <- read_csv(here::here("data/final/201709-hubway-tripdata.csv"))
bike_rider_17_oct <- read_csv(here::here("data/final/201710-hubway-tripdata.csv"))
bike_rider_19_aug <- read_csv(here::here("data/final/201908-bluebikes-tripdata.csv"))
bike_rider_19_sep <- read_csv(here::here("data/final/201909-bluebikes-tripdata.csv"))
bike_rider_19_oct <- read_csv(here::here("data/final/201910-bluebikes-tripdata.csv"))
## Bike Station
bike_stop <- read_csv(here::here("data/final/current_bluebikes_stations.csv"))

# Demographic Data
##


# Other Data
## MBTA Boundary
mbta_core <-
```


## R Markdown


